---
title: "Prosodic-Encoding-Walkthrough"
author: "Daniel Rud"
date: "2024-08-21"
output: md_document
---

# Introduction

In this file, we discuss the relevant software used to apply the prosodic encoding algorithm to texts in Section 4.1 of *Context Tree Classification and Clustering*.  


## Required libraries and sourced files

One needs to install the spacyr package and run `spacy_install()` and `spacy_initialize()`.  A fair warning -- spacyR may give some problems downloading, as there is often difficulty linking the package to a active python environment.  
```{r}
if (!requireNamespace("spacyr", quietly = TRUE)) install.packages("spacyr"); library(spacyr)
#Sys.setenv(SPACY_PYTHON = "~/myenvs/myenv/bin") # this is for my own virtual python environment for spacyr
spacy_install()
spacy_initialize()
if (!requireNamespace("tokenizers", quietly = TRUE)) install.packages("tokenizers"); library(tokenizers)
if (!requireNamespace("english", quietly = TRUE)) install.packages("english"); library(english)

source('DictionaryExtractionV2.R')
source('findNumSyl.R')
```

## Web Dictionary web-scraping 

A prerequisite to applying the Prosodic encoding algorithm is to know both the number of syllables and the location of a word's stressed syllable for all words in a text.  To do this, we originally implemented a web-scraping algorithm provided in the file *DictionaryExtractionV2.R*, where the functions `sAsENG` and `sAsBRIT` (sAs = stress AND syllables) were used to web-scrape the Merriam Webster and Cambridge dictionaries respectively.  However, due to recent changes in the web format of the Merriam Webster dictionary, I have augmented the function `sAsENG` to instead extract the United States English pronunciation from the Cambridge dictionary (which is provided).  The two functions `sAsENG` and `sAsBRIT` require that the user specify the part of speech a particular word pertains to.  In order to find the part of speech of a particular word in a sentence, we leverage the `spaCyr` package that uses NLP models to identify parts of speeches of tokenized words within sentences.  

For illustration, given the word "garage" and its part of speech "noun", the "stress AND syllables" can be found with the following call: 

```{r, eval = FALSE}
Word = "garage"
POS = "NOUN"

eng_SAS = sAsENG(Word = Word, type = POS)

brit_SAS = sAsBRIT(Word = Word, type = POS)

#OUTPUT: 

# [1] "Word: garage |  Entry: garage"
# [1] "numSyl: 2 |  Stress: 2"
# [1] "Word: garage |  Entry: garage"
# [1] "numSyl: 2 |  Stress: 1"
```
Notice that the word "garage" differs in the pronunciation between American and UK English.  


## Phonological Encoding Function 

We now provide the function to perform the phonological encoding.  The function is long; however, we will illustrate with an example in the chunk after the function initialization.  The phonological encoding process is described with detail in the manuscript *Context Tree Classification and Clustering*.  

```{r}
Phonological_Encoding = function(input_file, output_file = "", dictionary_type = c("Webster", "Cambridge"), elements = c("the", "a", "at")) {
  
  # Parse_Token function: Determines the prosodic encoding of a word based on its syllables, stress, and whether it begins a prosodic word
 Parse_Token <- function(Word, Num_Sil, Stressed_syllable, Output_file, Phonetic_word)
{       
    # Initialize an empty string to hold the prosodic encoding for the word
    prosodyString = ""
    
    # Proceed only if the word has one or more syllables
    if (Num_Sil > 0) # Ensure Num_Sil is positive
    {
        # Special case for words with exactly one syllable
        if (Num_Sil == 1)
        {	
            #### Check if the word is a non-stressed pronoun or determiner ####
            if (is.element(toupper(Word), c("ME", "US", "THEM", "HIM", "HER", "YOU", "IT", "A", "AN", "OF", "THE", "FOR"))) 
            {
                # Check if this is the beginning of a prosodic word
                if (Phonetic_word == 1) 
                { 
                    # Assign prosody code 2 (indicates start of a prosodic word)
                    prosodyString = "2 "
                    Phonetic_word <- 0 # Reset Phonetic_word flag
			    } 
                else 
                {
                    # Assign prosody code 0 (indicates middle of a prosodic word)
                    prosodyString = "0 "     			
     		    } 
     		} 
            # Case for one-syllable word that is not a non-stressed pronoun/determiner
            else
            {
                # Check if this is the beginning of a prosodic word
                if (Phonetic_word == 1) 
                { 
                    # Assign prosody code 3 (indicates stressed prosodic word)
                    prosodyString = "3 " 
			    } 
                else 
                {
                    # Assign prosody code 1 (stressed syllable in middle of prosodic word)
                    prosodyString = "1 " 
                    Phonetic_word = 1 # Update Phonetic_word to indicate this word was prosodic
     		    }	
     		}
     	} 
        # Case for multi-syllable words
     	else 
     	{
     		prosodyString = "" # Initialize/clear the prosody string
            # Loop through each syllable in the word
     		for (i in 1: Num_Sil)
     		{
                # If current syllable is stressed and it's the first syllable in the prosodic word
     			if ((Stressed_syllable == i) && (Phonetic_word == 1) && i == 1)
                {
                    # Assign prosody code 3 (stressed first syllable in prosodic word)
                    prosodyString = paste(prosodyString, "3 ", sep = "")
                    Phonetic_word = 0 # Reset Phonetic_word flag
                }
                # If current syllable is stressed but not the first syllable in the prosodic word
     			else if ((Stressed_syllable == i) && (Phonetic_word == 0))
                {
                    # Assign prosody code 1 (stressed syllable in middle of prosodic word)
     				prosodyString = paste(prosodyString, "1 ", sep = "")
                }
                # If current syllable is not stressed and it's the first syllable in the prosodic word
     			else if ((Stressed_syllable != i) && (Phonetic_word == 1))
     			{
                    # Assign prosody code 2 (unstressed first syllable in prosodic word)
     				prosodyString = paste(prosodyString, "2 ", sep = "")
                    Phonetic_word = 0 # Reset Phonetic_word flag
     			}
                # If current syllable is not stressed and not the first syllable in the prosodic word
				else if ((Stressed_syllable != i) && (Phonetic_word == 0))
                {
                    # Assign prosody code 0 (unstressed syllable in middle of prosodic word)
     				prosodyString = paste(prosodyString, "0 ", sep = "")
                }
     		}
     		Phonetic_word = 1 # Reset Phonetic_word flag for the next word
     	}	
    }
    
    # Write the prosody string to the output file and append it
    cat(prosodyString, file = Output_file, append = TRUE) 
    
    # Return the updated Phonetic_word flag and the prosody string
    return(c(Phonetic_word, prosodyString))
}

  
  ##################################################################################################
  ##################################################################################################
  ######## START HERE: Main body of the Phonological_Encoding function
  ##################################################################################################
  ##################################################################################################
  
  file <- input_file
  Output_file = 0
  
  # If no output file is provided, create a default output file name based on the input file name
  if(output_file == "") {
    Output_file <- paste(substr(file, 1, nchar(file) - 4), "_output.txt", sep = "")
    close(file(Output_file, open = "w")) # Clear the output file
  } else {
    Output_file = output_file
  }
  
  # Create a log file for debugging and tracking processing
  Output_log <- paste(substr(file, 1, nchar(file) - 4), "_log.txt", sep = "")
  
  # Read the entire input file
  text <- readChar(file, file.info(file)$size)
  
  # Tokenize text into sentences
  sentences = tokenize_sentences(text) 
  sink(Output_log)
  
  # Convert numbers to words in the text
  for (i in 1: length(sentences[[1]])) {
    sentence = unlist(tokenize_words(sentences[[1]][i]))
    sentence = gsub(",", "", sentence)
    suppressWarnings(numbers <- sapply(lapply(sentence, as.numeric), is.na))
    indecies = which(!numbers)
    
    # Replace numeric values with their text representation
    if (length(indecies) > 0) {
      for (j in indecies) {
        sentence[j] = toString(as.english(as.numeric(sentence[j])))
      }
    }
    
    # Rebuild sentences with converted numbers
    sentences[[1]][i] = toString(paste(paste(sentence, collapse = ' '), ".", sep = ""))
  }
  
  ################################################################################   
  
  # Process each sentence to compute prosodic encodings
  for (i in 1:length(sentences[[1]])) {
    
    Phonetic_word = 1 # Initialize flag for tracking prosodic words
    tokenize = spacy_parse(sentences[[1]][i], pos = TRUE) # Tokenize sentence into words and POS tags
    
    for (j in 1: dim(tokenize)[1]) { # Loop through each word in the sentence
      word = tokenize$token[j]
      pos = tokenize$pos[j]
      
      # Skip punctuation and certain contractions
      if (pos != "PUNCT" && pos != "SYM" && !is.element(tolower(word), c("'s", "’s", "n't", "n’t", "’t", "'t", "'ll", "’ll", "’d", "'d", "’n", "'n", "’ve", "'ve", "’re", "'re"))) {
        
        sink(Output_log, type = "output", split = TRUE, append = TRUE)
        
        # Default syllable and stress information
        sAs = 0
        
        # Retrieve syllable and stress information based on dictionary type
        if (dictionary_type == "Webster") {
          sAs = sAsENG(word, pos) # U.S. dictionary
        } else {
          sAs = sAsBRIT(word, pos) # British dictionary (if used)
        }
        
        sink()
        
        # If the word is not found in the dictionary, attempt to handle suffixes and prefixes
        if (is.na(sAs[1])) {
          result = WordAppendeges(word) # Analyze suffixes and prefixes
          Num_Sil <- as.numeric(result[3])
          Word <- result[2]
          fullWord = result[1]
          
          # Handle silent 'e' at the end of words
          Word = separateSilentE(Word, eDictionary)
          
          # Count syllables in the word
          for (k in 1: length(Word))
            Num_Sil = Num_Sil + findNumSil(Word[k])
          
          if (Num_Sil < 1)
            Num_Sil = 1
          
          sAs[1] = Num_Sil
          sAs[2] = 1
          
          sink(Output_log, append = TRUE, type = "output", split = TRUE)
          print("Word was not found in dictionary.  Revert to original function")
          print(paste("Word:", result[1]))
          print(paste("numSyl:", Num_Sil,  "|  Stress:", 1))
          sink()
        }
        
      }
      
      # Special case for abbreviations
      if (!is.na(sAs[1]) && sAs[1] == "abb") {
        sink(Output_log, append = TRUE, type = "output", split = TRUE)
        print(paste("Word:", word, "entered as abbreviation"))
        sink()
        
        # Each letter in the abbreviation gets a prosodic encoding of 3
        for (k in 1: nchar(word)) {
          cat("3 ", file = Output_file, append = TRUE)
        }
        
        Phonetic_word = 1 # Reset for next prosodic word
        
        sink(Output_log, type = "output", split = TRUE, append = TRUE)
        print(paste("Prosodic encoding for '", word, "' is: ", gsub(",", "", toString(rep("3 ", nchar(word)))), ".  Phonetic Word= ", Phonetic_word, sep = ""))
        print("--------------------------------------------------------------------")
        sink()
        
      } else if (pos != "PUNCT" && pos != "SYM" && !is.element(tolower(word), c("'s", "’s", "n't", "n’t", "’t", "'t", "'ll", "’s"))) {
        
        # Parse the token and assign prosodic encoding
        prosody = Parse_Token(word, sAs[1], sAs[2], Output_file, Phonetic_word)
        Phonetic_word = prosody[1]
        encoding = prosody[2]
        
        sink(Output_log, type = "output", split = TRUE, append = TRUE)
        print(paste("Prosodic encoding for '", word, "' is: ", encoding, ".  Phonetic Word= ", Phonetic_word, sep = ""))
        print("--------------------------------------------------------------------")
        sink()
        
      } else if (is.element(word, c(".", "?", "!"))) {
        
        # Sentence-ending punctuation gets a special encoding of 4
        cat("4 ", file = Output_file, append = TRUE)
        Phonetic_word = 1 # Reset for next prosodic word
        
        sink(Output_log, type = "output", split = TRUE, append = TRUE)
        print(paste("Prosodic encoding for '", word, "' is: 4.  Phonetic Word= ", Phonetic_word, sep = ""))
        print("--------------------------------------------------------------------")
        sink()
        
      }
    }
  }
}    
```

To showcase the function, let us first create a txt file containing two sentences.  

```{r}
text = "Police are hunting for answers on what caused an explosion that injured twenty nine people in New York's Chelsea neighborhood shortly before a second suspicious device was found nearby."
write(text, file = "sample_text.txt")
```

Now, let us apply the prosodic encoding function 

```{r, eval=FALSE}

suppressWarnings(Phonological_Encoding(input_file = "sample_text.txt", 
                      output_file = "sample_text_prosodic_output.txt", 
                      dictionary_type = "Webster"))

```

After running the Phonological encoding process, we will see that two files will be generated -- the Output file with the phonological text encoding, and the output log that includes details from the encoding process.   

Let us look at the output from the two files. 

The `sample_text_prosodic_output.txt` file will contain the phonological encoding of the text.  

```{r}
suppressWarnings(output_file_text <- readLines("sample_text_prosodic_output.txt"))

output_file_text
``` 


The `sample_text_log.txt` file will contain the log of the number of syllables and stress of each word, and the corresponding phonological encoding.  

```{r}
output_log_file_text = readLines("sample_text_log.txt")

output_log_file_text
```

## Full disclaimer 

The original algorithms developed to extract the phonetic spellings of words from Merriam-Webster and Cambridge dictionary webpages have been updated in response to recent modifications to these sites. Initially implemented in 2018, the original code has since become outdated as the structure and format of the webpages have evolved over time. Consequently, the revised algorithms **have not** undergone extensive validation for web scraping and may be prone to errors. In contrast, the original version successfully processed hundreds of text documents with minimal issues.

## Proposal for future prosodic encoding 

For future implementations of a prosodic encoding algorithm in text analysis, it may be beneficial to first explore existing literature for recurrent neural network architectures capable of predicting both syllable count and stress patterns, contingent on the language's country of origin. Given that we previously utilized spacyR for part-of-speech tagging, it is plausible that a comparable architectural approach could be leveraged to extract the desired phonological information, which we had initially attempted to acquire via web scraping










